Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-07-27 Paper/app.tex
..

\section{Application to $H_2$/$O_2$ reaction mechanism: high-dimensional case}
\label{sec:app}

>>>>

\section{$\text{H}_2$/$\text{O}_2$ reaction kinetics: high-dimensional case}
\label{sec:app}

For the high-dimensional case, we aim to investigate the impact of the uncertain
pre-exponents ($A_i$'s) as well as the activation energies ($E_{a,i}$'s) on ignition 
delay during the H$_2$/O$_2$ reaction. The $A_i$'s are considered to be uniformly
distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$. The $E_{a,i}$'s for all
reactions except $\mathcal{R}_6$ -- $\mathcal{R}_9$ and $\mathcal{R}_{13}$
are considered to be uncertain and uniformly distributed in the interval: 
$[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$. The nominal values, $A_i^\ast$ and $E_{a,i}^\ast$
corresponding to the different reaction rates are provided in~\cite{Yetter:1991}.

The grad-free approach discussed earlier in~\ref{sub:gradfree} was used to compute the
active subspace. Using the iterative procedure, the convergence of the eigenvectors
was examined by tracking $\delta w_{1,j}^{(i)}$. In Figure~\ref{fig:conv_app} (left),
components of the converged eigenvector based on 50 samples using $\tau$ = 0.2 are plotted.
To illustrate convergence, we compare them with components of the eigenvector obtained
using 100 samples (twice as many). The variation of maximum $\delta w_{1,j}^{(i)}$ between
successive iterations during the course of the iterative procedure is illustrated in 
Figure~\ref{fig:conv_app} (right). 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.8\textwidth]{./Figures/eigv10}
\caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector obtained
using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  $\delta w_{1,j}^{(r)}$
is plotted for successive iterations to illustrate the convergence behavior.}
\label{fig:conv_app}
\end{center}
\end{figure}
%

### 2018-07-27 Paper/intro.tex
..

We are interested in investigating the impact of uncertainty in the rate-controlling parameters, 
pre-exponent~($A_i$) and the activation energy~($E_{a,i}$) on ignition delay. The $A_i$'s are
considered to be uniformly distributed in the interval: $[0.9A_i^\ast, 1.1A_i^\ast]$.
The $E_{a,i}$'s for all reactions except $\mathcal{R}_6$ -- $\mathcal{R}_9$ and $\mathcal{R}_{13}$
are considered to be uncertain and uniformly distributed in the interval: $[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$.
$A_i^\ast$ and $E_{a,i}^\ast$ are nominal estimates corresponding to the $i^\text{th}$ reaction
as provided in~\cite{Yetter:1991}. Hence, the total number of uncertain parameters is 33 which makes the
present problem reasonably high-dimensional involving a significantly large amount of computational
effort. To reduce the effort, we explore two strategies: one involving the use of gradients (grad-based)
and the other involving local linear approximation which does require gradient computation (grad-free).
These strategies are based on algorithms 1.1 and 1.2 provided  in~\cite{Constantine:2015}. The amount of 
computational effort associated with the grad-free approach is significantly smaller than the grad-based
approach as discussed later in section~\ref{sec:method}.

>>>>

We are interested in investigating the impact of uncertainty in the rate-controlling parameters, 
pre-exponent~($A_i$) and the activation energy~($E_{a,i}$) on ignition delay. The $A_i$'s are
considered to be uniformly distributed about their nominal estimates provided in~\cite{Yetter:1991}.
The total number of uncertain inputs is 33 which makes the
present problem reasonably high-dimensional involving a significantly large amount of computational
effort. To reduce the effort, we explore two strategies: one involving the use of gradients (grad-based)
and the other involving local linear approximation which does require gradient computation (grad-free).
These strategies are based on algorithms 1.1 and 1.2 provided  in~\cite{Constantine:2015}. However, it
must be noted that in this work, an iterative approach is adopted for implementing the two strategies.
Consequently, the amount of computational effort appears to have decreased, and is found to be
relatively less in the case of grad-free approach as discussed later in section~\ref{sec:method}.
